---
title: "PM566_Assignment 04"
author: Mona Bandov
format:
  html:
    embed-resources: true
editor: visual
date: "2023-11-17"
---

# **Assignment 04 - HPC and SQL**

November 17, 2022

## **Due Date**

This assignment is due by midnight Pacific Time, November 17th, 2023.

The learning objectives are to write faster code for computational task requiring a loop and to implement some queries and basic data wrangling in SQL.

## **HPC**

### **Make things run faster**

Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at StackOverflow and Google

```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  # YOUR CODE HERE
  rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  # YOUR CODE HERE
  apply(mat,1,cumsum)
}

```

### **Question 1**

Using the dataset generated below (`dat`), check that the output of both of your new functions matches the output of the original functions. Then use `microbenchmark` to check that your version is actually faster.

```{r}
# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), unit = "relative"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), unit = "relative"
)
```

### **Make things run faster with parallel computing**

The following function allows simulating pi:

```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:

```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

### **Question 2**

Rewrite the previous code using `parLapply()` (or your parallelization method of choice) to parallelize it. Run the code once, using `system.time()`, to show that your version is faster.

```{r}
# YOUR CODE HERE
system.time({
  # YOUR CODE HERE
  ans <- # YOUR CODE HERE
  print(mean(ans))
  # YOUR CODE HERE
})


#FIX THIS CODE HERE 
```

## **SQL**

Setup a temporary database by running the following chunk

```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)

head(film)
head(category)
```

When you write a new chunk, remember to replace the `r` with `sql, connection=con`. Some of these questions will require you to use an inner join. Read more about them here <https://www.w3schools.com/sql/sql_join_inner.asp>

## **Question 3**

How many many movies are available in each `rating` category?

```{r}

movies_in_each_category <- dbGetQuery(con, "
  SELECT *
  FROM film
  INNER JOIN film_category ON film.film_id = film_category.film_id
")

head(movies_in_each_category)
```

## **Question 4**

What is the average replacement cost and rental rate for each `rating` category?

```{r}

avg_replacement <- dbGetQuery(con, "
  SELECT
    rating,
    AVG(replacement_cost) AS avg_replacement_cost,
    AVG(rental_rate) AS avg_rental_rate
  FROM
    film
  GROUP BY
    rating
")
head(avg_replacement)


```

## **Question 5**

Use table `film_category` together with `film` to find how many films there are with each category ID.

```{r}

film_in_each_category <- dbGetQuery(con, "
  SELECT
    fc.category_id,
    COUNT(f.film_id) AS film_count
  FROM
    film_category fc
  INNER JOIN
    film f ON fc.film_id = f.film_id
  GROUP BY
    fc.category_id
")
head(film_in_each_category)


```

## **Question 6**

Incorporate the `category` table into the answer to the previous question to find the name of the most popular category.

```{r}


most_popular_category <- dbGetQuery(con, "
  SELECT
    c.category_id,
    c.name AS category_name,
    COUNT(f.film_id) AS film_count
  FROM
    film_category fc
  INNER JOIN
    film f ON fc.film_id = f.film_id
  INNER JOIN
    category c ON fc.category_id = c.category_id
  GROUP BY
    c.category_id, c.name
  ORDER BY
    film_count DESC
  LIMIT 1
")

head(most_popular_category)



```
